Mapping mutational effects along the evolutionary landscape of HIV envelope

### 2018-02-15 paper/ResponseToReviews.tex

ModuleNotFoundError: No module named 'dms_tools2.protstruct'
\end{verbatim}

{\color{red}
This is now fixed by the latest version.}

7. It might be helpful to increase the transparency on the points in Figure 3B, because at the current level it's impossible to resolve any detail in the blob near zero.

{\color{black}
We have added a new figure supplement (Figure 3--Figure supplement 2) that shows the correlations as a KDE contour plot rather than a scatter plot.
As is apparent from this new supplement, the vast majority of points are near the origin (low preference in both replicates). 
This is expected since most mutations are highly deleterious.
Therefore, the correlation signal is due to relatively small fraction of amino acids that have non-negligible preference---a fact that makes sesne, since most mutations are so deleterious as to be evolutionarily irrelevant.
Unfortunately, given the skew of the data it is not possible to find a transparency that strikes a good balance between showing the high-preference points clearly and not over-plotting the low preference points.
Therefore, we have elected to keep the original Figure 3B (scatter plot) while also adding the new contour plot in Figure 3--Figure supplement 2) to show both elements of the data.
}

### 2018-02-15 paper/manuscript.tex
added contour version of correlation scatter plots

\begin{figure}
\begin{minipage}[t]{0.34\textwidth}
{\bf \Large A} \\ 
\centerline{{\includegraphics[clip=true, trim=4.3in 1.4in 0in 0.48in, width=0.4\textwidth]{figures/BG505_avgmutfreqs.pdf}}} 
\centerline{\small BG505}
\includegraphics[clip=true, trim=0in 0in 1.7in 0in, width=\textwidth]{figures/BG505_avgmutfreqs.pdf}
\centerline{\small BF520}
\includegraphics[clip=true, trim=0in 0in 1.7in 0in, width=0.89\textwidth]{figures/BF520_avgmutfreqs.pdf}
\end{minipage}
\begin{minipage}[t]{0.01\textwidth}
\end{minipage}
\begin{minipage}[t]{0.65\textwidth}
{\bf \Large B} \\
\includegraphics[width=\textwidth]{figures/allprefscorr.pdf}
\end{minipage}
\caption{\label{fig:mutfreqs}
The deep mutational scanning selects for functional Envs and yields measurements that are well correlated among replicates.
{\bf (A)}
The average per-codon mutation frequency when sequencing plasmids encoding wildtype Env (\emph{DNA}), plasmid mutant libraries (\emph{mutDNA}), mutant viruses after the final infection (\emph{mutvirus}), and virus generated from wildtype plasmids (\emph{virus}).
Mutations are categorized as nonsynonymous, synonymous, or stop codon. 
The \emph{DNA} samples show that sequencing errors are rare, and the \emph{virus} samples show that viral-replication errors are well below the frequency of mutations in the \emph{mutDNA} samples.
Comparing the \emph{mutvirus} to \emph{mutDNA} shows clear purifying selection against stop codons and some nonsynonymous mutations, particularly after subtracting the background error rates given by the \emph{virus} and \emph{DNA} samples (\FIGSUPP[mutfreqs]{mutfreqs}).
More extensive plots from the analysis of the deep sequencing data are in Supplementary~files~\ref{suppfile:code} and \ref{suppfile:html}.
{\bf (B)}
Correlations between replicates in the measured preferences of each site in Env for all 20 amino acids.
Blue indicates replicate measurements on BF520, red indicates replicate measurements on BG505, and gray indicates across-Env measurements of BF520 versus BG505.
$R$ is the Pearson correlation coefficient. 
The numerical values for the preferences are in \FIG{mutfreqs}-source~data~\ref{figdata:allprefs}.
}
\figsupp[\label{figsupp:mutfreqs} Numerical frequencies of mutations.]
{This table gives the average frequencies of nonsynonymous, synonymous, and stop-codon mutations as plotted in \FIG{mutfreqs}.
Note that there is only one \emph{DNA} sample for BF520, so that same sample is repeated three times in the sample in conjunction with each BF520 replicate.
We calculate the error-corrected \emph{pre}-selection mutation frequency as the \emph{mutDNA} frequency minus the \emph{DNA} frequency, and the error-corrected \emph{post}-selection mutation frequency as the \emph{mutvirus} frequency minus the \emph{virus} frequency.
We then use these error-corrected frequencies to calculate the \emph{percent} of mutations remaining after selection for each replicate and type of mutation.}
{\input{figures/avgmutfreqs}}
\figdata{\label{figdata:allprefs} Preferences for each replicate and averages are in \texttt{all\_prefs\_unscaled.zip}.}
\end{figure}

>>>>

\begin{figure}
\begin{minipage}[t]{0.34\textwidth}
{\bf \Large A} \\ 
\centerline{{\includegraphics[clip=true, trim=4.3in 1.4in 0in 0.48in, width=0.4\textwidth]{figures/BG505_avgmutfreqs.pdf}}} 
\centerline{\small BG505}
\includegraphics[clip=true, trim=0in 0in 1.7in 0in, width=\textwidth]{figures/BG505_avgmutfreqs.pdf}
\centerline{\small BF520}
\includegraphics[clip=true, trim=0in 0in 1.7in 0in, width=0.89\textwidth]{figures/BF520_avgmutfreqs.pdf}
\end{minipage}
\begin{minipage}[t]{0.01\textwidth}
\end{minipage}
\begin{minipage}[t]{0.65\textwidth}
{\bf \Large B} \\
\includegraphics[width=\textwidth]{figures/allprefscorr.pdf}
\end{minipage}
\caption{\label{fig:mutfreqs}
The deep mutational scanning selects for functional Envs and yields measurements that are well correlated among replicates.
{\bf (A)}
The average per-codon mutation frequency when sequencing plasmids encoding wildtype Env (\emph{DNA}), plasmid mutant libraries (\emph{mutDNA}), mutant viruses after the final infection (\emph{mutvirus}), and virus generated from wildtype plasmids (\emph{virus}).
Mutations are categorized as nonsynonymous, synonymous, or stop codon. 
The \emph{DNA} samples show that sequencing errors are rare, and the \emph{virus} samples show that viral-replication errors are well below the frequency of mutations in the \emph{mutDNA} samples.
Comparing the \emph{mutvirus} to \emph{mutDNA} shows clear purifying selection against stop codons and some nonsynonymous mutations, particularly after subtracting the background error rates given by the \emph{virus} and \emph{DNA} samples (\FIGSUPP[mutfreqs]{mutfreqs}).
More extensive plots from the analysis of the deep sequencing data are in Supplementary~files~\ref{suppfile:code} and \ref{suppfile:html}.
{\bf (B)}
Correlations between replicates in the measured preferences of each site in Env for all 20 amino acids.
Blue indicates replicate measurements on BF520, red indicates replicate measurements on BG505, and gray indicates across-Env measurements of BF520 versus BG505.
$R$ is the Pearson correlation coefficient. 
The numerical values for the preferences are in \FIG{mutfreqs}-source~data~\ref{figdata:allprefs}.
\FIGSUPP[mutfreqs]{contour} shows the correlations using contour rather than scatter plots.
}
\figsupp[\label{figsupp:mutfreqs} Numerical frequencies of mutations.]
{This table gives the average frequencies of nonsynonymous, synonymous, and stop-codon mutations as plotted in \FIG{mutfreqs}.
Note that there is only one \emph{DNA} sample for BF520, so that same sample is repeated three times in the sample in conjunction with each BF520 replicate.
We calculate the error-corrected \emph{pre}-selection mutation frequency as the \emph{mutDNA} frequency minus the \emph{DNA} frequency, and the error-corrected \emph{post}-selection mutation frequency as the \emph{mutvirus} frequency minus the \emph{virus} frequency.
We then use these error-corrected frequencies to calculate the \emph{percent} of mutations remaining after selection for each replicate and type of mutation.}
{\input{figures/avgmutfreqs}}
\figsupp[\label{figsupp:contour} Correlations plotted on a contour rather than a scatter plot.]
{
This figure shows the same data as in \FIG{mutfreqs}B except that it shows KDE contour plots rather than points.
As is obvious from this representation, the vast majority of the points fall near the origin (very low preference in both replicates), and most of the correlation signal is therefore due to the relatively modest number of amino acids that have high preference at any given site.
This is expected, since most amino acid mutations at most sites will be strongly deleterious, and so have low preference.
}{\includegraphics[width=\textwidth]{figures/allprefscorr_contour.pdf}}
\figdata{\label{figdata:allprefs} Preferences for each replicate and averages are in \texttt{all\_prefs\_unscaled.zip}.}
\end{figure}
